---
title: "bandersnatch"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(jsonlite)
library(tidyverse)
```

# Parse input data

```{r}
b <- readLines("bandersnatch.json") %>% fromJSON
```

```{r}
# Find decision point names from metadata section
choice_points <- b$interactiveVideoMoments$choicePointNavigatorMetadata$choicePointsMetadata$choicePoints
lapply(names(choice_points), function (choice_id) {
  choice <- choice_points[[choice_id]]
  tibble(id = choice_id, text = choice$description,
         choices =  list(choice$choices))
}) %>%
  bind_rows() ->
  choice_points
choice_points
```

## Moments

The main data object of the file - scenes defined by their time in the video and the choices presented.

```{r}
moments <- b$interactiveVideoMoments$momentsBySegment
# find all "moments" (scenes) and their available choices
moments_choices <- lapply(names(moments), function (moment_id) {
  moment <- jsonlite::flatten(moments[[moment_id]])
  # playbackImpression scene types alter the persistent game state.
  # They also have a time associated, but its not clear if they
  # actually play or if it just skips to the actual scene. 
  # Only tracking which scenes alter the saved state
  alters_state <- any(moment$type == "notification:playbackImpression")
  scenes <- grepl("scene:", moment$type)
  # skip "fakeOption" scenes (each has another real scene attached)
  if (!is.null(moment$trackingInfo.optionType)) {
    scenes <- scenes & ifelse(
      is.na(moment$trackingInfo.optionType),
      "",
      moment$trackingInfo.optionType
    ) != "fakeOption"
  }
  # Some ids have only a playbackImpression - not clear where
  # the flow goes from there
  if (!any(scenes)) {
    return(tibble(
      decision_id = moment_id, start_time = moment$startMs[1],
      end_time = moment$endMs[1], scene_type = moment$type[1],
      alters_state = alters_state
    ))
  }
  # After filtering impressions and fakes, max of 1 scene per id
  moment <- moment[scenes, ]
  choices <- moment$choices[[1]]
  if (moment$type == "scene:cs_bs_phone") {
    choices$text <- ifelse(
      is.na(choices$code),
      "Incorrect Code",
      choices$code
    )
  }
  choices <- choices[, intersect(names(choices), c("id", "segmentId", "sg", "text"))]
    
  choices %>% mutate(
    decision_id = moment_id,
    start_time = moment$startMs,
    end_time = moment$endMs,
    scene_type = moment$type,
    alters_state = alters_state
  )
  # choices_clean
}) %>%
  bind_rows()
moments_choices
```

## segment groups

Some choices lead to a segment group, which applies logic
from the game state to choose one of several options for the next
scene

```{r}

sgroups <- b$interactiveVideoMoments$segmentGroups
# data format varies depending on whether additional prereqs
# are defined for the group and whether some items link to other
# groups or just directly to scenes
doList <- function (x) {
  if (is.character(x)) {
    return (x)
  }
  if (!is.null(x$segment)) {
    return(x$segment)
  }
  if (is.list(x)) {
    return(doList(x))
  }
}
sgroups <- lapply(names(sgroups), function (group_id) {
  group <- sgroups[[group_id]]
  if (is.character(group)) {
    return (tibble(group_id = group_id, scene_id = group))
  }
  if (is.data.frame(group)) {
    tab <- tibble(group_id = group_id, scene_id = group$segment)
    tab$sub_group_id = group$segmentGroup
    return(tab)
  }
  return(tibble(group_id = group_id, scene_id = sapply(group, doList)))
}) %>%
  bind_rows()


sgroups <- sgroups %>% filter(!is.na(scene_id)) %>%
  select(-sub_group_id) %>%
  distinct() %>%
  rename(segmentId = scene_id)
sgroups
```

```{r}
# misnomer here - originally I was expanding out the segment groups, 
# but their recursive nature made that difficult and it works better
# visually to include them as "state branch" nodes in the network
moments_choices_unfurled <- moments_choices %>%
  distinct() %>%
  mutate(
    segmentId = ifelse(is.na(sg), segmentId, sg),
    id = ifelse(is.na(id), id, paste0("choice_", id))
  ) %>%
  select(-sg)
moments_choices_unfurled
```
# Prepare nodes

```{r}
# basic 2-way decision point scenes
dec_nodes <- moments_choices_unfurled %>%
  filter(!is.na(id)) %>%
  mutate(type = ifelse(
    scene_type == "scene:interstitialPostPlay_v2",
    "Reset Point",
    "Decision Point"
  )) %>%
  select(type, id = decision_id, start_time, end_time, alters_state) %>%
  distinct() %>%
  left_join(select(choice_points, id, text))
# segment groups as game state-based branch points
group_nodes <- sgroups %>%
  select(id = group_id) %>%
  distinct() %>%
  mutate(type = "State Branch")
# Nodes for the choices with their text label and destination
choice_nodes <- moments_choices_unfurled %>%
  filter(!is.na(id)) %>%
  select(id, text) %>%
  distinct() %>%
  mutate(type = "Choice")
nodes <- bind_rows(dec_nodes, group_nodes, choice_nodes)

# Miscellaneous & mysterious scenes or choices

# Scenes with no defined choices including credits scenes,
# but also others that appear to be dead-ends
nodes <- moments_choices_unfurled %>%
  filter(!decision_id %in% nodes$id) %>%
  select(id = decision_id,  start_time, end_time, alters_state) %>%
  distinct() %>%
  mutate(type = "Other Scene") %>%
  bind_rows(nodes, .)
# These scenes are referenced as destinations of choices, but not
# defined in moments or groups - unclear what they are
nodes <- moments_choices_unfurled %>%
  filter(!is.na(segmentId),
         !segmentId %in% nodes$id) %>%
  select(id = segmentId) %>%
  distinct() %>%
  mutate(type = "Other Scene") %>%
  bind_rows(nodes, .)
# These scenes are referenced as possible destinations from group
# branches, but not defined themselves in moments or groups 
nodes <- sgroups %>%
  filter(!segmentId %in% nodes$id) %>%
  select(id = segmentId) %>%
  distinct() %>%
  mutate(type = "Other Scene") %>%
  bind_rows(nodes, .)

nodes <- nodes %>% mutate(alters_state = factor(
  alters_state,
  levels = c(F, T, NA),
  labels = c("No", "Yes", "N/A"),
  exclude = NULL
)) 
  
nodes
```

# Prepare edges

```{r}
# Decision point -> choice options
dec_choice_edge <- moments_choices_unfurled %>%
  select(source = decision_id, target = id) %>%
  distinct() %>%
  mutate(edge_type = "decision-choice")
# Choice option -> next scene
choice_scene_edge <- moments_choices_unfurled %>%
  select(source = id, target = segmentId) %>%
  distinct() %>%
  mutate(edge_type = "choice-scene")
# Segment group -> all possible destinations
group_expand_edge <- sgroups %>%
  select(source = group_id, target = segmentId) %>%
  mutate(edge_type = "group")

edges <- bind_rows(dec_choice_edge, choice_scene_edge,
                   group_expand_edge) %>%
  filter(complete.cases(.))
edges
```

# Data validation

```{r}
# validation checks - all must have 0 records
filter(nodes, duplicated(id))
filter(edges, !edges$source %in% nodes$id)
filter(edges, !edges$target %in% nodes$id)
```

# Output

Output format is a ragged array of node and edge variables in long form. 

```{r}
c(
  rename(nodes, `Event Type` = type, `Game State Changed` = alters_state),
  edges
) %>%
  toJSON(na = "null", null = "null") %>%
  writeLines("bandersnatch Network.json")
```

